Conversations in time: interactive visualisation to explore structured temporal data

### 2020-08-11 tsibbletalk.Rmd
done with tidy temporal data review

The \CRANpkg{tsibble} package [@wang2020tsibble] introduces a unified temporal data structure, referred to as a `tsibble`, to represent time series and longitudinal data in a tidy format [@wickham2014tidy]. That said, a `tsibble` extends the `data.frame` and \CRANpkg{tibble} class with temporally contextual metadata: `index` and `key`. The `index` declares a data column that holds time-related indices. The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can contain multiple columns. Below displays the monthly Australian retail trade turnover data, available in the \CRANpkg{tsibbledata} package. The `Month` column holds year-months as `index`. The `State` together with `Industry` are the identifiers for these 152 series, highlighted as `key`. Note that the column `Series ID` could be an alternative option for setting up `key`, but `State` and `Industry` are more readable and informative. The `index` and `key` are "sticky" columns to a `tsibble`, forming critical pieces for fluent temporal data analysis later.

@@

In the spirit of tidy data to the \CRANpkg{tidyverse} [@Wickham2019], the **tidyverts** suite features `tsibble` as the foundational data structure, to build a fluid and fluent pipeline for time series analysis. Besides \CRANpkg{tsibble}, the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of statistical analysis and forecasting in the **tidyverts** ecosystem. When time series analysis starts taking off, series of interests defined by `key` are likely to remain unchanged over the course of analysis. Functions, such as `features()` and `model()`, summarise a sequence of indexed values down to a single statistic or model by every observational unit. The output is a normal table, where each row corresponds to an observational unit denoted by "key". In the context of relational databases, the "key" acts like a foreign key in a reduced form of tsibble, while the index and key together operates like a primary key. (tsibble-centered workflow)

@@

At the early stage of exploratory temporal data analysis, time series plots and scatterplots goes hand by hand. (insert figures below)

@@

The \pkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble data structure to enable reactivity for temporal data (i.e. a `tsibble`). The `as_shared_tsibble()` function turns a tsibble to a mutable tsibble `SharedTsibbleData`, as an R6 subclass of `SharedData` from {crosstalk}. A *tsibble* data object holds a set of interrelated series, identified by the "key", in a data frame. When the "key" includes more than two variables, there can exist intrinsic structures in all combinations of variables: either nesting or crossing. (Add an example here)

>>>>

The \CRANpkg{tsibble} package [@wang2020tsibble] introduces a unified temporal data structure, referred to as a `tsibble`, to represent time series and longitudinal data in a tidy format [@wickham2014tidy]. That said, a `tsibble` extends the `data.frame` and \CRANpkg{tibble} class with temporally contextual metadata: `index` and `key`. The `index` declares a data column that holds time-related indices. The `key` identifies a collection of related series or panels observed over the `index`-defined period, which can contain multiple columns. Below displays the monthly Australian retail trade turnover data (`aus_retail`), available in the \CRANpkg{tsibbledata} package. The `Month` column holds year-months as `index`. The `State` together with `Industry` are the identifiers for these 152 series, highlighted as `key`. Note that the column `Series ID` could be an alternative option for setting up `key`, but `State` and `Industry` are more readable and informative. The `index` and `key` are "sticky" columns to a `tsibble`, forming critical pieces for fluent temporal data analysis later.

@@

In the spirit of tidy data to the \CRANpkg{tidyverse} [@Wickham2019], the **tidyverts** suite features `tsibble` as the foundational data structure, in order to build a fluid and fluent pipeline for time series analysis. Besides \CRANpkg{tsibble}, the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of statistical analysis and forecasting in the **tidyverts** ecosystem. When time series analysis starts taking off, series of interest denoted by the `key` variables often remain unchanged over the course of analysis, from trend inspection to forecasting performance. Figure \ref{fig:highlight-retail} (a) gives an overview of 152 series for the retail data using an overlaid time series plot, while Figure \ref{fig:highlight-retail} (b) presents a scatterplot, where each series is represented by a dot in the feature space (trend versus seasonal strength). The plot making of Figure \ref{fig:highlight-retail} (b) is aided with the `features()` function from \CRANpkg{feasts}, which summarises original data by each series down to various statistical features. This function along with other **tidyverts** functions is `tsibble`-aware, and outputs a table in a reduced form where each row corresponds to a series, thus graphically displayed as Figure \ref{fig:highlight-retail} (b).

@@

Figure \ref{fig:highlight-retail} highlights not only a series with strongest seasonality, but also a need to querying interesting series on the fly. The two plots are static, but can be considered as linked views via the common `key` variables between two tables. This motivates enabling interactivity of `tsibble` and `tsibble`-derived objects for exploratory data analysis in a lightning manner.

@@

The \pkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk} package, introduces a shared tsibble data structure to enable reactivity for temporal data (i.e. a `tsibble`). The `as_shared_tsibble()` function turns a tsibble to a mutable tsibble `SharedTsibbleData`, as an R6 subclass of `SharedData` from {crosstalk}. A *tsibble* data object holds a set of interrelated series, identified by the "key", in a data frame. In the context of relational databases, the "key" acts like a foreign key in a reduced form of tsibble, while the index and key together operates like a primary key. (tsibble-centered workflow) When the "key" includes more than two variables, there can exist intrinsic structures in all combinations of variables: either nesting or crossing. (Add an example here)

### 2020-08-11 tsibbletalk.tex
done with tidy temporal data review

The \CRANpkg{tsibble} package \citep{wang2020tsibble} introduces a
unified temporal data structure, referred to as a \texttt{tsibble}, to
represent time series and longitudinal data in a tidy format
\citep{wickham2014tidy}. That said, a \texttt{tsibble} extends the
\texttt{data.frame} and \CRANpkg{tibble} class with temporally
contextual metadata: \texttt{index} and \texttt{key}. The \texttt{index}
declares a data column that holds time-related indices. The \texttt{key}
identifies a collection of related series or panels observed over the
\texttt{index}-defined period, which can contain multiple columns. Below
displays the monthly Australian retail trade turnover data, available in
the \CRANpkg{tsibbledata} package. The \texttt{Month} column holds
year-months as \texttt{index}. The \texttt{State} together with
\texttt{Industry} are the identifiers for these 152 series, highlighted
as \texttt{key}. Note that the column \texttt{Series\ ID} could be an
alternative option for setting up \texttt{key}, but \texttt{State} and
\texttt{Industry} are more readable and informative. The \texttt{index}
and \texttt{key} are ``sticky'' columns to a \texttt{tsibble}, forming
critical pieces for fluent temporal data analysis later.

@@

In the spirit of tidy data to the \CRANpkg{tidyverse}
\citep{Wickham2019}, the \textbf{tidyverts} suite features
\texttt{tsibble} as the foundational data structure, to build a fluid
and fluent pipeline for time series analysis. Besides \CRANpkg{tsibble},
the \CRANpkg{feasts} and \CRANpkg{fable} packages fill the role of
statistical analysis and forecasting in the \textbf{tidyverts}
ecosystem. When time series analysis starts taking off, series of
interests defined by \texttt{key} are likely to remain unchanged over
the course of analysis. Functions, such as \texttt{features()} and
\texttt{model()}, summarise a sequence of indexed values down to a
single statistic or model by every observational unit. The output is a
normal table, where each row corresponds to an observational unit
denoted by ``key''. In the context of relational databases, the ``key''
acts like a foreign key in a reduced form of tsibble, while the index
and key together operates like a primary key. (tsibble-centered
workflow)

@@

At the early stage of exploratory temporal data analysis, time series
plots and scatterplots goes hand by hand. (insert figures below)

@@

The \pkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk}
package, introduces a shared tsibble data structure to enable reactivity
for temporal data (i.e.~a \texttt{tsibble}). The
\texttt{as\_shared\_tsibble()} function turns a tsibble to a mutable
tsibble \texttt{SharedTsibbleData}, as an R6 subclass of
\texttt{SharedData} from \{crosstalk\}. A \emph{tsibble} data object
holds a set of interrelated series, identified by the ``key'', in a data
frame. When the ``key'' includes more than two variables, there can
exist intrinsic structures in all combinations of variables: either
nesting or crossing. (Add an example here)

>>>>

The \CRANpkg{tsibble} package \citep{wang2020tsibble} introduces a
unified temporal data structure, referred to as a \texttt{tsibble}, to
represent time series and longitudinal data in a tidy format
\citep{wickham2014tidy}. That said, a \texttt{tsibble} extends the
\texttt{data.frame} and \CRANpkg{tibble} class with temporally
contextual metadata: \texttt{index} and \texttt{key}. The \texttt{index}
declares a data column that holds time-related indices. The \texttt{key}
identifies a collection of related series or panels observed over the
\texttt{index}-defined period, which can contain multiple columns. Below
displays the monthly Australian retail trade turnover data
(\texttt{aus\_retail}), available in the \CRANpkg{tsibbledata} package.
The \texttt{Month} column holds year-months as \texttt{index}. The
\texttt{State} together with \texttt{Industry} are the identifiers for
these 152 series, highlighted as \texttt{key}. Note that the column
\texttt{Series\ ID} could be an alternative option for setting up
\texttt{key}, but \texttt{State} and \texttt{Industry} are more readable
and informative. The \texttt{index} and \texttt{key} are ``sticky''
columns to a \texttt{tsibble}, forming critical pieces for fluent
temporal data analysis later.

@@

In the spirit of tidy data to the \CRANpkg{tidyverse}
\citep{Wickham2019}, the \textbf{tidyverts} suite features
\texttt{tsibble} as the foundational data structure, in order to build a
fluid and fluent pipeline for time series analysis. Besides
\CRANpkg{tsibble}, the \CRANpkg{feasts} and \CRANpkg{fable} packages
fill the role of statistical analysis and forecasting in the
\textbf{tidyverts} ecosystem. When time series analysis starts taking
off, series of interest denoted by the \texttt{key} variables often
remain unchanged over the course of analysis, from trend inspection to
forecasting performance. Figure \ref{fig:highlight-retail} (a) gives an
overview of 152 series for the retail data using an overlaid time series
plot, while Figure \ref{fig:highlight-retail} (b) presents a
scatterplot, where each series is represented by a dot in the feature
space (trend versus seasonal strength). The plot making of Figure
\ref{fig:highlight-retail} (b) is aided with the \texttt{features()}
function from \CRANpkg{feasts}, which summarises original data by each
series down to various statistical features. This function along with
other \textbf{tidyverts} functions is \texttt{tsibble}-aware, and
outputs a table in a reduced form where each row corresponds to a
series, thus graphically displayed as Figure \ref{fig:highlight-retail}
(b).

@@

Figure \ref{fig:highlight-retail} highlights not only a series with
strongest seasonality, but also a need to querying interesting series on
the fly. The two plots are static, but can be considered as linked views
via the common \texttt{key} variables between two tables. This motivates
enabling interactivity of \texttt{tsibble} and \texttt{tsibble}-derived
objects for exploratory data analysis in a lightning manner.

@@

The \pkg{tsibbletalk} package, inspired by the \CRANpkg{crosstalk}
package, introduces a shared tsibble data structure to enable reactivity
for temporal data (i.e.~a \texttt{tsibble}). The
\texttt{as\_shared\_tsibble()} function turns a tsibble to a mutable
tsibble \texttt{SharedTsibbleData}, as an R6 subclass of
\texttt{SharedData} from \{crosstalk\}. A \emph{tsibble} data object
holds a set of interrelated series, identified by the ``key'', in a data
frame. In the context of relational databases, the ``key'' acts like a
foreign key in a reduced form of tsibble, while the index and key
together operates like a primary key. (tsibble-centered workflow) When
the ``key'' includes more than two variables, there can exist intrinsic
structures in all combinations of variables: either nesting or crossing.
(Add an example here)
